Capturing regeneration in models of forest dynamics

### 2022-11-09 manuscript/manuscriptRW.Rmd
added figures of species richness and corrected some details in the other figures

---
title: "Capturing regeneration in models of forest dynamics"
author:
- Olalla Díaz-Yáñez, Forest Ecology ETH Zurich
- Yannek Käber, Forest Ecology ETH Zurich
- Mats Mahnken, PIK Potsdam
- Samuel Fischer, UFZ
- Rico Fischer, UFZ
- Friedrich Bohn, UFZ
- Xavier Morin, cefe cnrs
- Manfred Lexer, BOKU
- Christian Hochauer, BOKU
- Katarina Merganicova, forim
- Tobias Mette, LWF Bayern
- Werner Rammer, TUM
- Kristin H. Braziunas, TUM
- Paola Mairota, UNIBA
- Josef Bruna, ibot cas
- Heike Lischke, WSL
- Daniel Scherrer, nWSL
- Simon Scheiter, senckenberg
- Thomas Hickler, senckenberg
- Tim Anders, senckenberg
- Jessica Hetzer, senckenberg
- Harald Bugmann, Forest Ecology, ETH Zurich
date: "`r Sys.Date()`"
output:
  word_document:
    reference_docx: template_paper.docx
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Forest recruitment is fundamental for forest dynamics, in particular
with increasing disturbance under climate change and it is also crucial
understand the forest resilience under climate change.

@@

Tree recruitment depends on several processes (reproduction, fruit
maturation, seed production, dispersal, germination, juvenile growth and
survival) that are all difficult to assess

Regeneration processes in dynamic forest models

Consequences of over recruit

Species richness in regeneration, mortality and regeneration levels
impacts on forest dynamics

This is the first time, a large number of forest models commonly used to
assess forest dynamics under climate change is evaluated against a
continental-scale multi-species harmonized data set on tree recruitment.

In this study wanted to answer the following questions:

1)  Are models of forests dynamics capturing correctly regeneration
    levels across environmental gradients

2)  Are models of forest dynamics representing correctly initial
    composition of tree populations

3)  Can models of forest dynamics correctly capture the initial
    mortality in tree in-growth between 7 and 10 cm

@@

````{r tableModels, echo=FALSE,  message=FALSE}
library(knitr)

@@

\begin{equation}
H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}(\#eq:shannon)
\end{equation}

\begin{equation}
p_{n,i} = \frac{BA_{r_{n,i}}}{BA_{total_{n}}}(\#eq:piShannon)
\end{equation}

where, $s$ is the total number of species in the sample $n$ (11 for the models that simulated all the species and a varied number for those models that did not consider all the species (see Table @ref(tab:tableModels) for more details in the simulated species); $p_{i}$ is the proportion of the species $i$ calculated as the recruited basal area ($BA_{r_{n,i}}$) for that species $i$ over the total basal area ($BA_{total_{n}}$) of that sample $n$.

@@

## Regeneration levels

-   Recruitment was overestimated across models both for 7 and 10 cm
    diameter thresholds, with the exception of SYBILA and Landis II that
    underestimated recruitment (Figure \@ref(fig:over)). The models with
    higher overestimation are PICUS and TreeMig. The variability across
    sites in recruitment levels is similar or smaller to the observed
    variability for most of the models, with the exception of PICUS and
    TreeMig where the mean recruitment variability is larger.

    ```{r over, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites. The red dotted lines shows the 2.5 and 97.5 percentiles for the 7cm diameter threshold", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
    ```

@@

```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites. The values have been divided in 10 bins and the red lines represent a GAM model showing the trend from the observed data.", out.width='100%'}
    knitr::include_graphics(here::here("figures", "envTrend7Totba.png"))
```

```{r nvTrend710BA500, echo=FALSE, fig.cap="BA to be join with previous figure", out.width='100%'}
    knitr::include_graphics(here::here("figures", "envTrend7Totba_500.png"))
```

@@

````{=html}
<!--The relationship was first described by
[@key] for single citation
[@key1; @key2] multiple citation can be separated by semi-colon
[-@key] in order to suppress author name, and just display the year
[see @key1 p 12; also this ref @key2] is alos a valid syntax
```
-->
````

>>>>

---
title: "Capturing regeneration in models of forest dynamics"
author:
- Olalla Díaz-Yáñez, Forest Ecology ETH Zurich
- Yannek Käber, Forest Ecology ETH Zurich
- Mats Mahnken, PIK Potsdam
- Samuel Fischer, UFZ
- Rico Fischer, UFZ
- Friedrich Bohn, UFZ
- Xavier Morin, cefe cnrs
- Manfred Lexer, BOKU
- Christian Hochauer, BOKU
- Katarina Merganicova, forim
- Tobias Mette, LWF Bayern
- Werner Rammer, TUM
- Kristin H. Braziunas, TUM
- Paola Mairota, UNIBA
- Josef Bruna, ibot cas
- Heike Lischke, WSL
- Daniel Scherrer, nWSL
- Simon Scheiter, senckenberg
- Thomas Hickler, senckenberg
- Tim Anders, senckenberg
- Jessica Hetzer, senckenberg
- Harald Bugmann, Forest Ecology, ETH Zurich
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
  word_document:
    reference_docx: template_paper.docx
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: draft.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Tree recruitment is fundamental for forest dynamics, in particular with
increasing disturbance under climate change and it is also crucial
understand the forest resilience under climate change.

@@

Tree recruitment depends on multiple processes (reproduction, fruit
maturation, seed production, dispersal, germination, juvenile growth and
survival) that are all difficult to assess

Models are needed to (1) synthesize existing empirical data and explore
their consequences, and (2) assess future tree recruitment e.g. in the
context of global change scenarios

Regeneration processes in dynamic forest models and how they have been
handled to date [@konig_tree_2022; @bugmann2022].

Tendency of models to be too liberal with recruitment; consequences of
over-recruitment Species richness in regeneration, mortality and
regeneration levels may have strong impacts on forest dynamics

An analysis how well models of forest dynamics perform re. recruitment
is lacking. This is the first time, a large number of forest models
commonly used to assess forest dynamics under climate change is
evaluated against a continental-scale multi-species harmonized data set
on tree recruitment. Briefly praise the dataset [@käber].

In this study wanted to answer the following questions: 1) Are models of
forests dynamics capturing correctly regeneration levels across
environmental gradients 2) Are models of forest dynamics representing
correctly initial composition of tree populations 3) Can models of
forest dynamics correctly capture the initial mortality in tree
in-growth between 7 and 10 cm

@@

```{r tableModels, echo=FALSE,  message=FALSE}
library(knitr)

@@

```{=tex}
\begin{equation}
H_n = - \sum_{i = 1}^{s}{p_{i}ln(p_{i})}(\#eq:shannon)
\end{equation}
```
```{=tex}
\begin{equation}
p_{n,i} = \frac{BA_{r_{n,i}}}{BA_{total_{n}}}(\#eq:piShannon)
\end{equation}
```
where, $s$ is the total number of species in the sample $n$ (11 for the
models that simulated all the species and a varied number for those
models that did not consider all the species (see Table
@ref(tab:tableModels) for more details in the simulated species);
$p_{i}$ is the proportion of the species $i$ calculated as the recruited
basal area ($BA_{r_{n,i}}$) for that species $i$ over the total basal
area ($BA_{total_{n}}$) of that sample $n$.

@@

Regeneration levels, species composition and mortality between
recruitment diameters thresholds of 7 and 10 cm varied across model
simulation results and in comparison with the observations. As expected,
recruitment was overestimated across most models for both the 7 and 10
cm diameter thresholds, with the exception of SYBILA, Landis II and
aDGVM2 that estimated recruitment rates at the lower end of the
plausibility interval of the empirical data (Figure \@ref(fig:over)).
The models with highest overestimation were PICUS and TreeMig. For most
of the models, the variability across sites in recruitment levels
(visible from the IQR
in the box plots of Figure \@ref(fig:over)) is similar to or smaller
than the observed
variability for most of the models, with the exception of PICUS, where
the mean
recruitment variability is much larger.

```{r over, echo=FALSE, fig.cap="Mean recruitment values across all samples for the 200 sites. The red dotted lines shows the 2.5 and 97.5 percentiles for the 7 cm diameter threshold.", out.width='100%'}
        knitr::include_graphics(here::here("figures", "recOverAll710.png"))
```

@@

```{r nvTrend710BA, echo=FALSE, fig.cap="Mean recruitment values across samples for the 200 sites. The values have been divided in 10 bins and the red lines represent a GAM model showing the trend from the observed data.", fig.show = 'hold', out.width = '50%'}
    knitr::include_graphics(c(here::here("figures", "envTrend7Totba.png"),
                              here::here("figures", "envTrend7Totba_500.png")))
```

@@

```{=html}
<!--The relationship was first described by
[@key] for single citation
[@key1; @key2] multiple citation can be separated by semi-colon
[-@key] in order to suppress author name, and just display the year
[see @key1 p 12; also this ref @key2] is alos a valid syntax
-->
```
\`\`\`
